### 2017-05-10 paper/evaluation.tex
pass on eval

\section{Evaluation}
\label{sec:eval}

@@

\newcommand{\numRQA}{RQ1}
\newcommand{\numRQB}{RQ2}
\newcommand{\numRQD}{RQ3}
\newcommand{\numRQC}{RQ4}
\newcommand{\numRQE}{RQ5}
\newcommand{\numRQF}{RQ6}

\newcommand{\RQA}{How prevalent is the occurence of time-consuming
  test suites\Comment{ in open-source projects}?}
\newcommand{\RQB}{How time is distributed across test cases?}
\newcommand{\RQC}{How prevalent is the use of test suite
  parallelization\Comment{ in open-source projects}?}
\newcommand{\RQD}{What are the speedups obtained with parallelization
  (in projects that actually use it) on cost?}
\newcommand{\RQE}{What are the main reasons that prevent developers
  from using test suite parallelization?}
\newcommand{\RQF}{To which extent parallel execution configurations affect testing
  costs and flakiness?}

\begin{itemize}
\item Potential
  \begin{itemize}
  \item \textbf{\numRQA.} \RQA
  \item \textbf{\numRQB.} \RQB    
  \end{itemize}
\item Effectiveness
  \begin{itemize}
  \item \textbf{\numRQD.} \RQD
  \end{itemize}  
\item Adoption
  \begin{itemize}
  \item \textbf{\numRQC.} \RQC    
  \item \textbf{\numRQE.} \RQE
  \end{itemize}
\item Tradeoffs
  \begin{itemize}
  \item \textbf{\numRQF.} \RQF    
  \end{itemize}          
\end{itemize}

@@

\subsection{Potential}
\label{sec:rqA}
\label{sec:rqB}

@@

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  It is therefore important to
speedup regressing testing in open-source projects.\Comment{not only
  to huge projects as those from Google~\cite{google-tap,google-ci}
  and Microsoft~\cite{prasad-shulte-ieee-microsoft-ci}.}

Research question \numRQB{} assesses whether parallelism can help in
those cases.  One important factor to the effectiveness of parallelism
is the distribution of test costs in the test suite.  In the limit, if
cost is dominated by a single test from a large test suite, it is
unlikely that parallelization will be beneficial as a test method is
the smallest working unit in test frameworks.  However, avoiding
frequent context switches is another factor to consider.  For example,
assuming there are at least two CPUs available for execution, cost can
be cut in half if two tests in a large test suite dominate execution
time. For the case where cost is distributed more evenly across test
cases, one expects that speedups will be a function of the number of
cores.

%% These contradictory forces, pushing number of tests and cost
%% of each test up and down, make prediction of effectiveness challenging.

@@

\sloppy Figures~\ref{fig:longtcost} and~\ref{fig:medtcost} show the
time distribution of individual test cases per project.  We observed
that the average median value of execution cost for a test was
relatively small (dashed horizontal red lines), namely 0.31s for
\medg{} projects and 0.23s for \longg{} projects.  The standard
deviations associated with each distribution were relatively
low.\Comment{ Figure~\ref{fig:sd} shows the number of projects within
  specific ranges of $\sigma$ values.}  We noted a small number of
cases of CPU monopolization.  For example, the highest value of
$\sigma$ occurred in \CodeIn{uber\_chaperone}, a project from the
medium group.  This project contains only 65 tests, 62 of which take
less than 0.5s to run, one of which takes nearly 3s to run, and two of
which take $\sim$40m to run.  For this project, 99.2\% of the
execution cost is dominated by only 3\% of the tests; without these
two costly tests this project would have been classified as
short-running.  A closer inspection in the data indicates that the
project \CodeIn{uber\_chaperone} was a corner case: we did not find
projects with such extreme time monopolization profile.  Project
\CodeIn{facebookarchive\_linkbench} is also classified as long-running
and has the second highest value of $\sigma$.  For this project,
however, cost is distributed more smoothly across \Fix{529} tests, of
which \Fix{119 (23\%)} take more than \Fix{1s} to run with the rest of
the tests running faster.

@@

Figure~\ref{fig:time-distributions} showed that the average median
times were similar.  A closer inspection on results indicates that the
difference in overall running times across projects in these groups is
mainly justified by the number of test cases as opposed to the
individual costs of test cases.  Figure~\ref{fig:size-testsuites}
shows the difference in the distribution of test suite sizes across
groups.  This figure indicates that long projects, albeit having a
wider inter-quartile range (middle 50\% projects in this group are
less predictable), have a higher median and much higher averages.
Furthermore, we noted a strong positive correlation between running
time and number of test on projects in the \longg{} group.  The
positive correlation between these two variables was, however, weak in
the \medg{} group, suggesting that saving time in this group with test
suite parallelization may be more challenging as relatively fewer
tests dominate overall execution time.  Figure~\ref{fig:scattercost}
shows these results.

@@

\subsection{Effectiveness}
\label{sec:rqD}

\begin{figure}[h!]
\centering
\resizebox{.48\textwidth}{!}{%
\begin{tabular}{llrrl}
\toprule
\emph{Group} & \emph{Subject} & \multicolumn{1}{c}{$T_s$} & \multicolumn{1}{c}{$T_p$} & $T_s/T_p$ \\%
\midrule%
Medium & BounceStorage \textbf{Chaos} HTTP Proxy  & 1.51m & 1.47m & 
    \cellcolor{lightgray}1.02x\\%
Medium & Apache \textbf{Flink} & 11.79m & 2.57m & 4.59x\\%
Long & Jenkins CI \textbf{Gerrit} Plugin & 51.19m & 40.31m &  1.26x\\%
Medium & Spotify \textbf{Helios} & 4.46m & 1.63m & 2.73x\\%  
Medium & \textbf{Javaslang} & 2.18m & 1.82m & 1.19x\\%
Medium & \textbf{Jcabi} GitHub & 2.76m & 0.30m &
    \cellcolor{lightgray}9.2x\\%
Medium & Hazelcast \textbf{Jet} & 8.26m & 3.67m & 2.25x\\%
Medium & Apache \textbf{Log4J2} & 8.24m & 8.21m & 
    \cellcolor{lightgray}1.00x\\%
Medium & Jankotek \textbf{MapDB} & 10.06m & 8.58m & 1.17x\\%
\midrule
\textbf{average} &  &  &  & \avgSpeedup{}x\\
\bottomrule%
\end{tabular}}
\caption{Speedup (or slowdown) of parallel execution ($T_p$) over
  sequential execution ($T_s$).  Default parallel configuration of
  Maven is used.  Abbreviation of project names appear in bold face
  and highest slowdown/speedup appear in gray color.}
\label{tab:speedup}
\end{figure}

@@

To answer this research question, we conducted an experiment involving
the six parallel configurations from Section~\ref{sec:modes}  and the
top 10 long-running test modules from our sample set. The rationale
for this selection criteria was to maximize the chances of observing
speedup and flakiness, assuming that long-running tests also have many
tests. Indeed, we confirmed that test suites in these projects contain
at least 236 tests ($1,662.7$ in average). Also, we preferred to compare the
effect of a configuration over a single test module for projects with
multiple test modules.  Recall that large projects may contain several
test modules, and these modules may contain distinct characteristics
that could favor one configuration and not others; therefore, it would
be necessary to check each module individually for a fair comparison.
We are interested in understanding how efficiency (\ie, testing cost)
and flakiness (\ie, failing tests) are affected when we run a test
suite with different parallel configurations.  Recall that increased
resource contention obtained with parallelism can lead to concurrency
issues such as data races.  Flakiness and speedup are contradictory
forces that drive configuration selection.  We used as a \emph{control
group} (\ie, baseline) the sequential execution of each subject's
tests. Notice that for measuring flakiness, we have to consider only
tests that failed due to the concurrency in the parallel execution.
For that reason, we re-executed the tests in sequence ten times and
carefully verified that there are only passing tests in our baseline.
We considered ten as the number of re-executions based on the approach
used by Google reported in a previous work on test
flakiness~\cite{luo-etal-fse2014}.  To obtain parallel configurations,
we implemented a script that takes a subject and a configuration (\eg,
\ForkSeq{}) as inputs, and the script outputs the modified version of
the subject with the desired configuration. The workflow consists in
copying the project directory to a new directory, finding all existing
build files (\ie, \pomf{} files), and modifying all existing Maven
Surefire configurations with new values for the parameters
\CodeIn{parallel}, \CodeIn{forkCount}, and \CodeIn{threadCount} using
an XPath~\cite{xpath} library to manipulate XML documents. For
configurations with forked JVMs enabled (\ie, \ForkSeq{} and
\ForkParMeth{}), we changed the \CodeIn{forkCount} with the value
\CodeIn{1C} (\ie, one JVM per core).  To adjust the pool of threads
for parallelism within a JVM, we changed the parameter
\CodeIn{threadCount} with \Jbc{should we consider 2 as it is the
number of native threads per core OR 6 as it represents 3 Cores * 2
native threads?}.  To run the subjects and their respective variations
(\ie, the modified versions according to the parallelism
configuration), we used a similar approach as described in
Figure~\ref{fig:mvn-execution} except that we added a timeout of one
hour to run the tests. We used the \CodeIn{timeout}
command~\cite{timeout-cmd} to monitor the execution, and we configured
the command to dispatch a \emph{kill} signal if the test execution
exceeds the time limit. We imposed this time constraint to avoid
hanging indefinitely the experiment execution due to some thread
contention that may occur (\eg, deadlock). Finally, we saved each
execution log and XML test reports generated by Maven to collect the
execution time, the number of failing tests, and for reference to
analyze and diagnose outliers in our results. For efficiency, we
reported the speedup (\ie, $\Uparrow_\text{speed} = T_{\text{s}} /
T_{\text{p}}$) in average, and for flakiness, we reported the rate of
failing tests (\ie, $\mathit{\%_\text{fail} = failures / tests}$) in
average.  Figure~\ref{tab:rq6-table} summarizes the obtained results
ordered by the time cost (\ie, $T_\text{cost}$). \Fix{Remember to
explain efficiency cutoff, JVM crashes and timeout}

@@

@@

>>>>

\newpage
\section{Evaluation}
\label{sec:eval}

@@

\newcommand{\numRQA}{RQ1}
\newcommand{\numRQB}{RQ2}
\newcommand{\numRQC}{RQ3}
\newcommand{\numRQE}{RQ4}
\newcommand{\numRQF}{RQ5}
\newcommand{\numRQD}{RQ6}
\newcommand{\numRQG}{RQ7}

\newcommand{\RQA}{How prevalent is the occurence of time-consuming
  test suites\Comment{ in open-source projects}?}
\newcommand{\RQB}{How time is distributed across test cases?}
\newcommand{\RQC}{How popular is test suite
  parallelization\Comment{ in open-source projects}?}
\newcommand{\RQD}{What are the speedups obtained with parallelization
  (in projects that actually use it) on cost?}
\newcommand{\RQE}{What are the main reasons that prevent developers
  from using test suite parallelization?}
\newcommand{\RQF}{How parallel execution configurations affect testing
  costs and flakiness?}
\newcommand{\RQG}{How test execution scales with the number of
  available CPUs?}

\setlist[itemize]{leftmargin=1em}
\begin{itemize}
\item Feasibility
  \begin{itemize}
  \item \textbf{\numRQA.} \RQA
  \item \textbf{\numRQB.} \RQB    
  \end{itemize}
\item Adoption
  \begin{itemize}
  \item \textbf{\numRQC.} \RQC    
  \item \textbf{\numRQE.} \RQE
  \end{itemize}
\item Tradeoffs
  \begin{itemize}
  \item \textbf{\numRQF.} \RQF    
  \end{itemize}
\item Speedups
  \begin{itemize}
  \item \textbf{\numRQD.} \RQD
  \item \textbf{\numRQG.} \RQG
  \end{itemize}    
\end{itemize}

@@

\subsection{Feasibility}
\label{sec:rqA}
\label{sec:rqB}

@@

Section~\ref{sec:rqA} showed that medium and long-running projects are
not uncommon, accounting to nearly \percentMedLongRunning{} of the
\numSubjs{} projects we analyzed.  Research question \numRQB{}
measures the distribution of test costs in test suites as to estimate
(lack of) potential of obtaining speedups with parallelization.  In
the limit, if cost is dominated by a single test from a large test
suite, it is unlikely that parallelization will be beneficial as a
test method is the smallest working unit in test frameworks.

%% However, avoiding
%% frequent context switches is another factor to consider.  For example,
%% assuming there are at least two CPUs available for execution, cost can
%% be cut in half if two tests in a large test suite dominate execution
%% time and these tests are assigned to different CPUs.

%% It is therefore important to
%% speedup regressing testing in open-source projects.\Comment{not only
%%   to huge projects as those from Google~\cite{google-tap,google-ci}
%%   and Microsoft~\cite{prasad-shulte-ieee-microsoft-ci}.}

%% For the case
%% where cost is distributed more evenly across test cases, one expects
%% that speedups will be a function of the number of cores.
%% These contradictory forces, pushing number of tests and cost
%% of each test up and down, make prediction of effectiveness challenging.

@@

\sloppy Figures~\ref{fig:longtcost} and~\ref{fig:medtcost} show the
time distribution of individual test cases per project.  We observed
that the average median value of execution cost for a test was
relatively small (dashed horizontal red lines), namely 0.31s for
\medg{} projects and 0.23s for \longg{} projects.  The standard
deviations associated with each distribution were relatively
low.\Comment{ Figure~\ref{fig:sd} shows the number of projects within
  specific ranges of $\sigma$ values.}  We noted a small number of
cases of CPU monopolization.  For example, the highest value of
$\sigma$ occurred in \CodeIn{uber\_chaperone}, a project from the
medium group.  This project contains only 65 tests, 62 of which take
less than 0.5s to run, one of which takes nearly 3s to run, and two of
which take $\sim$40m to run.  For this project, 99.2\% of the
execution cost is dominated by only 3\% of the tests; without these
two costly tests this project would have been classified as
short-running.  A closer inspection in the data indicates that the
project \CodeIn{uber\_chaperone} was a corner case; we did not find
other projects with such extreme time monopolization profile.  Project
\CodeIn{facebookarchive\_linkbench} is also classified as long-running
and has the second highest value of $\sigma$.  For this project,
however, cost is distributed more smoothly across \Fix{529} tests, of
which \Fix{119 (23\%)} take more than \Fix{1s} to run with the rest of
the tests running faster.

@@

Figure~\ref{fig:time-distributions} showed that the average median
times were similar for \medg{} and \longg{}-running test suites.
Results indicate that the difference in overall running times of
projects in those groups was mainly justified by the number of test
cases as opposed to the individual costs of test cases.
Figure~\ref{fig:size-testsuites} shows the difference in the
distribution of test suite sizes across groups.  This figure indicates
that long projects, albeit having a wider inter-quartile range (middle
50\% projects in this group are less predictable), have a higher
median and much higher average number of test cases.  Furthermore, we
noted a strong positive correlation between running time and number of
test on projects in the \longg{} group.  Considering the \medg{}
group, the correlation between these two variables was weak.
Figure~\ref{fig:scattercost} illustrates this correlation.

%% however, weak in
%% the suggesting that saving time in this group with test suite
%% parallelization may be more challenging as relatively fewer tests
%% dominate overall execution time.  Figure~\ref{fig:scattercost} shows
%% these results.

@@

@@

@@

To answer this research question, we conducted an experiment involving
the six parallel configurations from Section~\ref{sec:modes}  and the
top 10 long-running test modules from our sample set. The rationale
for this selection criteria was to maximize the chances of observing
speedup and flakiness, assuming that long-running tests also have many
tests. Indeed, we confirmed that test suites in these projects contain
at least 236 tests ($1,662.7$ in average). Also, we preferred to compare the
effect of a configuration over a single test module for projects with
multiple test modules.  Recall that large projects may contain several
test modules, and these modules may contain distinct characteristics
that could favor one configuration and not others; therefore, it would
be necessary to check each module individually for a fair comparison.
We are interested in understanding how efficiency (\ie, testing cost)
and flakiness (\ie, failing tests) are affected when we run a test
suite with different parallel configurations.  Recall that increased
resource contention obtained with parallelism can lead to concurrency
issues such as data races.  Flakiness and speedup are contradictory
forces that drive configuration selection.  We used as a \emph{control
group} (\ie, baseline) the sequential execution of each subject's
tests. Notice that for measuring flakiness, we have to consider only
tests that failed due to the concurrency in the parallel execution.
For that reason, we re-executed the tests in sequence ten times and
carefully verified that there are only passing tests in our baseline.
We considered ten as the number of re-executions based on the approach
used by Google reported in a previous work on test
flakiness~\cite{luo-etal-fse2014}.  To obtain parallel configurations,
we implemented a script that takes a subject and a configuration (\eg,
\ForkSeq{}) as inputs, and the script outputs the modified version of
the subject with the desired configuration. The workflow consists in
copying the project directory to a new directory, finding all existing
build files (\ie, \pomf{} files), and modifying all existing Maven
Surefire configurations with new values for the parameters
\CodeIn{parallel}, \CodeIn{forkCount}, and \CodeIn{threadCount} using
an XPath~\cite{xpath} library to manipulate XML documents. For
configurations with forked JVMs enabled (\ie, \ForkSeq{} and
\ForkParMeth{}), we changed the \CodeIn{forkCount} with the value
\CodeIn{1C} (\ie, one JVM per core).  To adjust the pool of threads
for parallelism within a JVM, we changed the parameter
\CodeIn{threadCount} with \Jbc{should we consider 2 as it is the
number of native threads per core OR 6 as it represents 3 Cores * 2
native threads?}.  To run the subjects and their respective variations
(\ie, the modified versions according to the parallelism
configuration), we used a similar approach as described in

Figure~\ref{fig:mvn-execution} except that we added a timeout of one
hour to run the tests. We used the \CodeIn{timeout}
command~\cite{timeout-cmd} to monitor the execution, and we configured
the command to dispatch a \emph{kill} signal if the test execution
exceeds the time limit. We imposed this time constraint to avoid
hanging indefinitely the experiment execution due to some thread
contention that may occur (\eg, deadlock). Finally, we saved each
execution log and XML test reports generated by Maven to collect the
execution time, the number of failing tests, and for reference to
analyze and diagnose outliers in our results. For efficiency, we
reported the speedup (\ie, $\Uparrow_\text{speed} = T_{\text{s}} /
T_{\text{p}}$) in average, and for flakiness, we reported the rate of
failing tests (\ie, $\mathit{\%_\text{fail} = failures / tests}$) in
average.  Figure~\ref{tab:rq6-table} summarizes the obtained results
ordered by the time cost (\ie, $T_\text{cost}$). \Fix{Remember to
explain efficiency cutoff, JVM crashes and timeout}

@@

\subsection{Speedups}
\label{sec:rqD}

\begin{figure}[h!]
\footnotesize
\centering
\resizebox{.45\textwidth}{!}{%
\begin{tabular}{llrrr}
\toprule
\emph{Group} & \emph{Subject} & \multicolumn{1}{c}{$T_s$} & \multicolumn{1}{c}{$T_p$} & $T_s/T_p$ \\%
\midrule%
Medium & \Comment{BounceStorage }Chaos\Comment{ HTTP Proxy} & 1.51m & 1.47m & 
    \cellcolor{lightgray}1.01x\\%
Medium &\Comment{ Apache }Flink& 11.79m & 2.57m & 4.59x\\%
Long &\Comment{ Jenkins CI }Gerrit\Comment{ Plugin} & 51.19m & 40.31m &  1.26x\\%
Medium &\Comment{ Spotify }Helios& 4.46m & 1.63m & 2.73x\\%  
Medium &Javaslang& 2.18m & 1.82m & 1.19x\\%
Medium &Jcabi\Comment{ GitHub} & 2.76m & 0.30m &
    \cellcolor{lightgray}9.2x\\%
Medium &\Comment{ Hazelcast }Jet& 8.26m & 3.67m & 2.25x\\%
Long &\Comment{ Apache }Log4J2& 8.24m & 8.21m & \cellcolor{lightgray}1.00x\\%
Long &\Comment{ Jankotek }MapDB& 10.06m & 8.58m & 1.17x\\%
\midrule
\textbf{average} &  &  &  & \avgSpeedup{}x\\
\bottomrule%
\end{tabular}}
\caption{Speedup (or slowdown) of parallel execution ($T_p$) over
  sequential execution ($T_s$).  Default parallel configuration of
  Maven is used.  Abbreviation of project names appear in bold face
  and highest slowdown/speedup appear in gray color.}
\label{tab:speedup}
\end{figure}
